Predicting Remaining Useful Life of NASA Aircraft Engines

### 2025-03-09 index.qmd
3/9 updates

@@

$H(t) = \int_0^t h(t) dt$
where,
-   $H(t)$ is the cumulative hazard at time $t$

@@

@@

### Data Exploration and Visualization

@@

@@

```{r, echo = FALSE}
#packages
library(tidyverse)
library(gridExtra)
library(survival)
library(survminer)
library(ggpubr)
library(magrittr)
library(ggfortify)
library(knitr)
library(sjPlot)

When using Cox proportional hazards models to model survival data, the
data must include a value for time, a unique identifier column
indicating which individual the time series is for, covariates used for
predictions, and an event indicating the status of the individual at
each point in time. The data used in this analysis contain all of the
previously mentioned fields with the exception of a status field.

@@

The visuals below describe the distribution of the 260 engines in the
training data. The histogram shows the survival times are normally
distributed with a median value of 199 iterations and the boxplot shows
there are few outliers out of the 260 engines. The engines have a median
survival time of 199 iterations with a standard deviation of 46.8. The
longest surviving engine lasted 378 iterations, with the shortest
survival time for all engines being 128.

@@

@@

### Modeling

Creating model

@@

**Checking Proportional Hazards**

>>>>

After verifying the survival data is formatted properly, the CPH model can be created in R. Important steps to consider are choosing covariates and verifying the model meet the model's assumptions after the model is created.

@@

$H(t) = \int_0^t h(t) dt$
where,

-   $H(t)$ is the cumulative hazard at time $t$

@@

Time until event can be calculated by modeling survival probability for future time values. When survival probability drops below 50%, it will be assumed the event occurred. Predictions for time until event will be made on the testing data for each individual's maximum time value. A column will be added to the table showing the expected time until event.

Tim until event can be modeled in R by passing the `type = "lp"` argument in the `stats::predict()` function to get a value for linear predictor.

@@

```{r, echo = FALSE, warning=FALSE,include=FALSE}
#Load packages
library(tidyverse)
library(gridExtra)
library(survival)
library(survminer)
library(ggpubr)
library(magrittr)
library(ggfortify)
library(knitr)
library(sjPlot)

### Data Structure, Exploration, and Visualization

@@

#### Data Structure

@@

@@

#### Data Visualization

The visuals below describe the distribution of the 260 engines in the
training data. The histogram shows the survival times are normally
distributed with a median value of 199 iterations and the box plot shows
there are few outliers out of the 260 engines. The engines have a median
survival time of 199 iterations with a standard deviation of 46.8. The
longest surviving engine lasted 378 iterations, with the shortest
survival time for all engines being 128.

@@

#### Choosing Covariates

@@

#### Creating Model

@@

#### Checking Assumptions

**Checking Proportional Hazards Assumption**
